---
title: "Codebook.rmd"
author: "Aditya Garapati, Carl Walenciak, Brian Waite"
date: "2/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(kableExtra)
library(readxl)
```

## DDS Analytics Talent Management Study Codebook

This codebook provides an overview of the data as it was received and any modifications that were required to support our analysis. It is a documentation of the variables in the data set, any normalization or modifications that were made, and should support replication of our analysis. 

### Data Files Received

DDSAnalytics provided unique(Consulting) with a Microsoft Excel (.xlsx) file titled CaseStudy2-data.xlsx. 

```{r}
cs2Raw <- read_excel("CaseStudy2-data.xlsx")
```

### Initial Data Structure

The data contained within the provided table has the following variable names and types: 
<br>

```{r}
kable(t(lapply(cs2Raw,class)))%>%
  kable_styling()%>%
  scroll_box(height = 100, width = 500)
```

Derived variables include: 

| Variable              | Values        | Purpose                                                |
|-----------------------|:-------------:|-------------------------------------------------------:|
| AttritionCode         | 1 or 0        | If Attrition = "Yes", AttritionCode = 1 and vice versa |
| OvertimeCode          | 1 or 0        | If Overtime = "Yes", OvertimeCode = 1 and vice versa   |
| maleperc              | integer       | Percentage of employees that are male from total       |
| femaleperc            | integer       | Percentage of employees that are female from total     |
| mean_hr               | integer       | Mean Job satisfaction of HR employees                  |
| mean_non_hr           | integer       | Mean Job satisfaction of Non-HR employees              |
| jobroleAttritionRates | data.frame    | A data frame containing attrition rates per job role   | 

### Test Models

A number of test models were created to help identify features provided in the data set that contribute to Attrition rates within the population. 

#### Model 1: Logistic Regression

A logistic regression model to predict whether an employee -- based on provided features -- would be in class (AttritionCode = 1) or out of Class. This model accepts all of the provided categorical and continuous variables and determines their impact on Attrition. The model is un-normalized and unweighted and provides only a first step analysis. Despite this, however, it is capable of achieving accuracy rates of 87% with Precision values of 90.7% and Recall of 94.9%. 

This model is described by the statement: 

```{r eval=FALSE, echo=TRUE}
AttritionCode ~ Age + JobRole + StockOptionLevel + WorkLifeBalance + 
              YearsSinceLastPromotion + EnvironmentSatisfaction + JobInvolvement + 
              OverTime + RelationshipSatisfaction + YearsAtCompany + YearsWithCurrManager + 
              BusinessTravel + DistanceFromHome + Gender + MaritalStatus + 
              NumCompaniesWorked + YearsInCurrentRole
```

