---
title: "Codebook.rmd"
author: "Aditya Garapati, Carl Walenciak, Brian Waite"
date: "2/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(kableExtra)
library(readxl)
```

## DDS Analytics Talent Management Study Codebook

This codebook provides an overview of the data as it was received and any modifications that were required to support our analysis. It is a documentation of the variables in the data set, any normalization or modifications that were made, and should support replication of our analysis. 

### Data Files Received

DDSAnalytics provided unique(Consulting) with a Microsoft Excel (.xlsx) file titled CaseStudy2-data.xlsx. 

```{r}
cs2Raw <- read_excel("CaseStudy2-data.xlsx")
```

### Initial Data Structure

The data contained within the provided table has the following variable names and types: 
<br>

```{r}
kable(t(lapply(cs2Raw,class)))%>%
  kable_styling()%>%
  scroll_box(height = 100, width = 500)
```




