---
title: "DDSAnalytics: Attrition and Job Role Study"
author: Carl Walenciak, Brian Waite, Aditya Garapati
output: html_document
---
<style type="text/css">
    
    h3 {
      text-align: center;
    }
    
    hr {
        border-color: #292F33;
        border-width: 1px;
    }

</style>




```{r setup, echo=F, include=FALSE}
library(manipulate)
library(plyr)
library(ggplot2)
library(scales)
library(knitr)
library(magrittr)
library(kableExtra)

df <- read.csv( "CaseStudy2Data.csv", stringsAsFactors=FALSE)

#Add Attrition Code
df$AttritionCode <- 0
df[df$Attrition == "Yes", ]$AttritionCode <- 1

#Add IsMale
df$IsMale <- 0
df[df$Gender == "Male", ]$IsMale <- 1

#Shorten Names
df[df$Department == "Research & Development", ]$Department <- "R&D"
df[df$Department == "Human Resources", ]$Department <- "HR"

#Age Grouping
df$AgeGroup <- ""
df[df$Age < 20, ]$AgeGroup <- "<20"
df[df$Age >= 20 & df$Age < 30, ]$AgeGroup <- "20-29"
df[df$Age >= 30 & df$Age < 40, ]$AgeGroup <- "30-39"
df[df$Age >= 40 & df$Age < 50, ]$AgeGroup <- "40-49"
df[df$Age >= 50 & df$Age < 60, ]$AgeGroup <- "50-59"
df[df$Age >= 60, ]$AgeGroup <- "60+"
df$AgeGroup <- factor(df$AgeGroup, level = c("<20", "20-29", "30-39", "40-49", "50-59", "60+"))
```
<br> 

## Introduction: 

Our team is pleased to submit the following analysis in response to DDSAnalytics request for support in identifying factors contributing to attrition and any other possible trends associated with job roles within the work force and provided data set. In the conduct of analysis, we have employed a number of exploratory data analysis techniques and are confident that we have unmasked some significant insights into these requests. 

<br> 

## Analysis Repository: 

All of the files, code, and presentation materials used in support of this submission are available to DDS Analytics at the following GitHub repository: 
https://github.com/Ujustwaite/DDS_Case_Study_2

<br> 

## Workforce Overview / Initial Statistics:

Being unfamiliar with the DDSAnalytics workforce or structure, we began our analysis with a robust examination of the operating construct, workforce composition, job roles, departments and other aspects contained within the data. 

The data we were provided contained records for *1,470* current and former employees of DDSAnalytics. These records broke down as follows. 

<!-------------------------------------------------->
<!-------------- DEPARMENT BREAKDOWN --------------->
<!-------------------------------------------------->

<br/>
<br/>

### **Department Breakdown**
<hr/>

<br/>  
We observed three departments within the organization. The Human Resources or HR departement was by far the smallest and Research and Development or R&D was the largest, with Sales somewhere in between. 
<br/> 

```{r department_break, fig.align='center',  echo=FALSE}
df_dep <- aggregate(df$EmployeeCount, by=list(df$Department), FUN=sum, drop=FALSE)
colnames(df_dep) <- c("Department", "Employee Count")
df_dep$`% of Total Employees` <- percent(df_dep$`Employee Count`/sum(df_dep$`Employee Count`))
```


```{r department_chart, fig.align='center', echo=FALSE}
blank_theme <- theme_minimal()+
    theme(
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.border = element_blank(),
        panel.grid=element_blank(),
        axis.ticks = element_blank(),
        plot.title=element_text(size=14, face="bold")
    )

bp <- (ggplot(df_dep, aes(x="", y=`Employee Count`, fill=Department))
        + geom_bar(width = 1, stat = "identity")
        + scale_fill_manual(values=c( "skyblue1",'cadetblue3','lightblue4'))
        + theme_minimal())

(bp + coord_polar("y", start=0)
    + blank_theme
    + theme(axis.text.x=element_blank(), legend.position="top")
)

```


```{r display_departments, fig.align='center', echo=FALSE, }
kable(df_dep, row.names = F, align=c('l',rep('c', length(df_dep) - 1) )) %>% kable_styling(bootstrap_options = "striped", full_width = F)
```

<br/>  
<br/>  

<!-------------------------------------------------->
<!------------------- AGE GROUP -------------------->
<!-------------------------------------------------->


### **Age Group**
<hr/>

<br/>  

The age distribution of the population represented by the data is normally distributed as we might expect. In general, we do not see any concerning trends here with regards to the age of the workforce. In other words, there is no massive wave of retirees impending nor are there any indications of a junior-biased workforce that would indicate a lack of professional experience. We examine this further, but at this point the data appear to be representative of the national population as a whole. 
<br/> 
```{r age_group_plot, fig.align='center', echo=F}

(ggplot(df, aes(x=AgeGroup, y = 1, fill=Gender)) 
    + geom_bar(stat="identity") 
    + scale_fill_manual(values=c('lightblue4','cadetblue3'))
    + theme_minimal()
    + xlab("Age Group")
    + ylab("Number of Employees")
    + theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)))
 )
```

<br/>

```{r age_group_table, fig.align='center', echo=FALSE}
emp_count <- aggregate(df$EmployeeCount, by=list(df$AgeGroup), FUN=sum, drop=FALSE)
colnames(emp_count) <- c("AgeGroup", "Employee Count")

male_count <- aggregate(df$IsMale, by=list(df$AgeGroup), FUN=sum, drop=FALSE)
colnames(male_count) <- c("AgeGroup", "Male Employees")

df2 <- merge(emp_count, male_count, by="AgeGroup")
df2$`Male Employee %` <- percent(df2$`Male Employees`/df2$`Employee Count`)

kable(df2, row.names = F, align=c('l',rep('c', length(df2) - 1) )) %>% kable_styling(bootstrap_options = "striped", full_width = F)
```

<br/>  
<br/>  

<!-------------------------------------------------->
<!------------------- JOB ROLE --------------------->
<!-------------------------------------------------->

### **Job Role**
<hr/>

<br/>  
Nine (9) unique job roles were identified within the data. Sales representatives, research scientists, and laboratory technicians represent the greatest number of positions. From the below plot we can see that there is a slight gender bias in the workforce. The true percentage breakdown is 60% male and 40% female, but that distribution varies greatly by job role. Human resources representatives, for example, are overwhelmingly male (69.2%). Female concerns may not be sufficiently represented in the HR department and consideration should be given to adding additional female HR reps. Efforts could be made to improve female representation in general, because only a small number of roles are near a 50-50 distribution. 
<br/> 
```{r job_role_plot, fig.align='center', echo=FALSE}

(ggplot(df, aes(x=JobRole, y =1, fill = Gender)) 
    + geom_bar(stat="identity") 
    + scale_fill_manual(values=c('lightblue4','cadetblue3'))
    + theme_minimal()
    + xlab("Job Role")
    + ylab("Number of Employees")
    + theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)))
    + theme(axis.text.x = element_text(angle = 60, hjust = 1))
)

```

<br/>  

```{r job_role_table, fig.align='center', echo=FALSE }
job_role <- aggregate(df$EmployeeCount, by=list( df$Department, df$JobRole), FUN=sum)
colnames(job_role) <- c("Department",  "Job Role", "Employee Count")
job_role$id <- paste(job_role$Department, job_role$`Job Role`)

job_male <- aggregate(df$IsMale, by=list(df$Department, df$JobRole), FUN=sum)
colnames(job_male) <- c("Department",  "Job Role", "Male Employees")
job_male$id <- paste(job_male$Department, job_male$`Job Role`)

df3 <- merge(job_role, job_male[, c("id", "Male Employees")], by="id")
df3$`Male Employee %` <- percent(df3$`Male Employees`/df3$`Employee Count`)
df3$id <- NULL

kable(df3, row.names = F, align=c('l','l', rep('c', length(df3) - 1) )) %>% kable_styling(bootstrap_options = "striped", full_width = F)
```

<br/>
<br/>


<!-------------------------------------------------->
<!---------------- JOB ROLE TRENDS ----------------->
<!-------------------------------------------------->

### **Job Role Trends**
<hr/>

<br/>


```{r job_role_trends, echo=FALSE}
dept_role <- aggregate(df$EmployeeCount, by=list(df$Department, df$JobRole), FUN=sum, drop=TRUE)
colnames(dept_role) <- c("Department", "Job Role", "Employee Count")
dept_role$id <- paste(dept_role$Department, dept_role$`Job Role`)

#Calc Percentage
dept_role$Percentage <- percent(dept_role$`Employee Count`/sum(dept_role$`Employee Count`))

colnames(dept_role) <- c("Department", "Job Role", "Employee Count","id", "% of Total Employees")
```


```{r attrition_merge, echo=FALSE}
df_attr <- df[df$AttritionCode == 1, ]
df_attr <- aggregate(df_attr$AttritionCode, by=list(df_attr$Department, df_attr$JobRole), FUN=sum, drop=FALSE)
df_attr[is.na(df_attr)] <- 0
colnames(df_attr) <- c("Department", "JobRole", "Attrition Count")
df_attr$id <- paste(df_attr$Department, df_attr$JobRole)

# Merge Combined and Yearst At Company
dept_role_comb <- merge(dept_role, df_attr[,c('id', 'Attrition Count')], by="id")

# Attrition Percentage
dept_role_comb$`Attrition Rate per Dept.` <- percent(dept_role_comb$`Attrition Count`/dept_role_comb$`Employee Count`)
```


```{r years_at_co_merge, echo=FALSE}
years_at_comp <- aggregate(df$YearsAtCompany, by=list(df$Department, df$JobRole), FUN=mean)
years_at_comp$id <- paste(years_at_comp$Group.1, years_at_comp$Group.2)
colnames(years_at_comp)[3] <- 'Avg. Years At Company'
years_at_comp$`Avg. Years At Company` <- round(years_at_comp$`Avg. Years At Company`, 1)

# Merge Combined and Yearst At Company
dept_role_comb <- merge(dept_role_comb, years_at_comp[,c('id', 'Avg. Years At Company')], by="id")
```


```{r job_sat_merge, echo=FALSE}
df_job_sat <- aggregate(df$JobSatisfaction, by=list(df$Department, df$JobRole), FUN=mean)
df_job_sat$id <- paste(df_job_sat$Group.1, df_job_sat$Group.2)

colnames(df_job_sat)[3] <- 'Avg. Job Satisfaction'
df_job_sat$`Avg. Job Satisfaction` <- round(df_job_sat$`Avg. Job Satisfaction`, 1)

# Merge
dept_role_comb <- merge(dept_role_comb, df_job_sat[,c('id', 'Avg. Job Satisfaction')], by="id")

```


```{r env_sat_merge, echo=FALSE}
env_sat <- aggregate(df$EnvironmentSatisfaction, by=list(df$Department, df$JobRole), FUN=mean)
env_sat$id <- paste(env_sat$Group.1, env_sat$Group.2)

colnames(env_sat)[3] <- 'Avg. Env. Satisfaction'
env_sat$`Avg. Env. Satisfaction` <- round(env_sat$`Avg. Env. Satisfaction`, 1)

# Merge
dept_role_comb <- merge(dept_role_comb, env_sat[,c('id', 'Avg. Env. Satisfaction')], by="id")

```


```{r display_combined_table, echo=FALSE}

dept_role_comb$id <- NULL

kable(dept_role_comb, row.names = F, align=c('l','l', rep('c', length(dept_role_comb) -2) ))%>% kable_styling(bootstrap_options = "striped", full_width = F)
```

<br/>
<br/>  

<!-------------------------------------------------->
<!---------- NUMBER OF COMPANIES WORKED ------------>
<!-------------------------------------------------->

### **Number of Companies Worked**
<hr/>

<br/>  

```{r exp_setup, echo=FALSE}
df$NumWorkGroup <- ""
df[df$NumCompaniesWorked < 2, ]$NumWorkGroup <- "<2"
df[df$NumCompaniesWorked >= 2 & df$NumCompaniesWorked < 5, ]$NumWorkGroup <- "2-4"
df[df$NumCompaniesWorked >= 5 & df$NumCompaniesWorked < 8, ]$NumWorkGroup <- "5-7"
df[df$NumCompaniesWorked >= 8, ]$NumWorkGroup <- "8+"
df$AgeGroup <- factor(df$AgeGroup, level = c("<2", "2-4", "5-7", "8+"))
```


```{r num_companies, echo=FALSE}
num_comp <- aggregate(df$EmployeeCount, by=list(df$NumWorkGroup), FUN=sum)
colnames(num_comp) <- c("# Companies Worked", "Employee Count")

num_comp_attr <- aggregate(df$AttritionCode, by=list(df$NumWorkGroup), FUN=sum)
colnames(num_comp_attr) <- c("# Companies Worked", "Attrition Count")

num_comb <- merge(num_comp, num_comp_attr, id=`# Companies Worked`)
num_comb$`Attrition Rate per Group` <- percent(num_comb$`Attrition Count`/num_comb$`Employee Count`)
```

<br/>

```{r num_companies_plot, fig.align='center', echo=FALSE}

(ggplot(df, aes(x=NumWorkGroup, y =1, fill =Attrition)) 
    + geom_bar(stat="identity") 
    + scale_fill_manual(values=c('lightblue4','cadetblue3'))
    + theme_minimal()
    + xlab("Number of Companies Worked")
    + ylab("Number of Employees")
    + theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)))
)

```

```{r num_companies_table, echo=FALSE}
kable(num_comb, row.names = F, align=c('l',rep('c', length(num_comb) - 1) ))%>% kable_styling(bootstrap_options = "striped", full_width = F)
```
